Optimal Decomposition-Recombination Plans for 2D Well-Constrained Graphs

### 2014-12-04 doc/GMP2015/s01_introduction.tex
s01_introduction.tex edited online with Bitbucket

\section{Introduction}
Modeling, analysis and design of self-similar, aperiodic layered materials using geometric constraint systems

Many natural and engineered materials are stacked two dimensional (2D) layers, where the structure within each layer may be recursively 
self-similar \cite{Intro1}, 
spanning multiple scales. Within any one scale, the material layer is aperiodic, and quasi-uniform,
consisting of a few repeated motifs appearing in disordered arrangments.
Each layer is not necessarily planar, i.e., it consists of multiple, inter-constraining planar (genus 0) monolayers. 
Furthermore, a layer is often   either isostatic or somewhat flexible, and rarely self-stressed, at any scale. These properties are
consistent with a self-assembled 2D structure that minimizes mass, optimally distributes external stresses and itself 
participates in the assembly of diverse and multifunctional, 
larger 2D structures.

@@

In order to study structural and mechanical properties such as rigidity, flexes and distribution of external stresses 
within the material layer, it is natural to model a material layer as a solution of geometric constraint systems 
arising from appropriate geometric primitives under metric or algebraic constraints. 
Such  2D {\em sesamacs} (Self-similar or aperiodic material constraint systems) can be used to 
to understand or design material layers (their solutions) with desired properties.

As constraint systems, 2D sesamacs  are  well-studied: so-called
{\it bar-joint} systems (Examples 1,2 above), {\it multibody-pin} systems (Example 4,5) or {\it pinned-line incidence} systems (Example 3).  
All of these systems have highly efficient combinatorial algorithms (pebble-game \cite{XX}, or equivalently, simplified network flow \cite{XX}) 
for locating maximal, generically rigid subsystems \cite{XX}.

Whether self-similar, or just aperiodic, recursive decomposition of 2D sesamacs 
is imperative for understanding or designing rigidity, flexes and distribution 
of external stresses as well as robustness/sensitivity to constraint variations -- at all scales. 
In fact, the Decomposition-Recombination (DR)
-planning problem  for a geometric constraint system has been formalized \cite{XX} and 
studied by many \cite{XX}. The problem is purely combinatorial  - since decomposition does not require
knowledge of the geometry. In addition, there are efficient  algorithms 
for optimal recombination of solutions to child systems, in order to obtain a solution to the parent system at any
given level of the DR-plan.

In particular, since many different such decompositions typically exist, is to ensure 
that any system is decomposed into a minimal set 
of maximal rigid proper subsystems  
such that they can be recombined efficiently 
 Two key observations serve as the motivations for this paper:

>>>>

\section{Introduction}
Modeling, analysis and design of quasi-uniform or self-similar material layers using geometric constraint systems

Many natural and engineered materials are stacked two dimensional (2D) layers, where the structure within each layer may be recursively 
self-similar \cite{Intro1}, 
spanning multiple scales. Within any one scale, the material layer is generally aperiodic, and quasi-uniform,
consisting of a few repeated motifs appearing in disordered arrangments.
Each layer is not necessarily planar, i.e., it consists of multiple, inter-constraining planar (genus 0) monolayers. 
Furthermore, a layer is often   either isostatic or somewhat flexible, and rarely self-stressed, at any scale. These properties are
consistent with a self-assembled 2D structure that minimizes mass, optimally distributes external stresses and itself 
participates in the assembly of diverse and multifunctional, 
larger 2D structures.

@@

In order to study structural and mechanical properties such as rigidity, flexes and distribution of external stresses 
within the material layer, it is natural to model a material layer as a solution of geometric constraint systems 
arising from appropriate geometric primitives under metric or algebraic constraints. 
Such  2D {\em qusecs} (quasi-uniform or self-similar constraint systems) can be used to 
to understand or design material layers (their solutions) with desired properties.

We now briefly survey existing techniques and their limitations, for 2D {\em qusecs,} which are most naturally treated as
{\it bar-joint} systems (Examples 1,2 above), {\it multibody-pin} systems (Example 4,5) or {\it pinned-line incidence} systems (Example 3). 
The limitations directly motivate the contributions of this paper.

First, all of these have highly efficient combinatorial algorithms (pebble-game \cite{XX}, or equivalently, simplified network flow \cite{XX}), 
for locating maximal, generically rigid subsystems \cite{XX}. When the input is rigid, these algorithms do nothing, i.e, the output is the same as the input.
However, both for self-similar or just aperiodic 2D qusecs, recursive decomposition into rigid subsystems
is imperative for understanding or designing rigidity, flexes, distribution 
of external stresses as well as robustness/sensitivity to constraint variations -- at all scales. 
In particular, recursive decomposition of rigid systems down to the geometric primitives is necessary.

Second,  decomposition of geometric constraint systems has been formalized \cite{XX} and well-studied \cite{XX} 
as the Decomposition-Recombination (DR)
-planning problem.
For the types of 2D qusecs listed above, generic rigidity is 
a combinatorial property and hence decomposition does not require
knowledge of the geometric information in the constraint system. 
In particular, since many different such decompositions typically exist, 
some criteria  to ensure good or optimal DR-plans were given in \cite{XX}. An optimal DR-plan is one
that minimizes the maximum number of child subsystems, over all its parent systems, a property that is 
crucial for obtaining a solution to the parent constraint system from the solutions of the child systems.
For special 2D qusecs such as so-called {\em tree-decomposable} systems, optimal DR-planning is immediate.

the optimal DR-planning problem was shown to be NP-hard for general 2D qusecs.

Other criteria where any parent system is composed of  
a minimal set of maximal rigid proper subsystems

Third, In addition, there are general algorithms for
recombination of  child system solutions, in order to obtain a solution to the parent system at any
given level of the DR-plan.

 Two key observations serve as the motivations for this paper:
